Quantum Magic Games

### 2023-07-04 paper/main.tex
paper => ex for first direction of main result

\section{Contextuality}
\label{sec:context}

@@

  \begin{figure}[H]
    \begin{center}
      \begin{tikzpicture}[vertex/.style={circle, thin, fill=blue!30, draw, inner sep=0, minimum size=5pt}]
        \node (00) [vertex] at (-1.25, 0) {};
        \node (01) [vertex] at (0,  0) {};
        \node (02) [vertex] at (1.25,  0) {};

@@

    A classical realization \(c\) for this arrangement is
  \begin{align*}
    c(v_{00}) = c(v_{01}) = c(v_{10}) = c(v_{11}) &= +1 \\
    c(v_{02}) = c(v_{12}) &= -1.
  \end{align*}
\end{tcbexample}

@@

  By repeatedly contracting edges, eventually we reach a graph with a
  single vertex.  Due to how we defined the label of each new vertex
  (i.e.\ as the product of the labels of the vertices being merged),
  it follows that the label of the last vertex equals the product of
  the labels of all the original vertices, and therefore equals the
  parity of the arrangement. The remaining vertex has multiple
  self-loops, two for each measurement \(M\) ascribed to the original
  edges.  Since \(M^{2} = I\), going counterclockwise we get a
  sequence \(M_{0}^{2} \ldots M_{\vert E'\vert -1}^{2} = I\). But, we
  saw that contraction preserves the third property, which says that
  the observables around a vertex multiply to the identity times the
  vertex label. Thus, the label of the final vertex is \(1\) and,
  consequently, the parity of the arrangement is \(+1\). \qed
\end{proof}
\printbibliography{}
\end{document}

>>>>

\subsection{Contextuality}
\label{sec:context}

@@

  \begin{figure}[H]
    \begin{center}
      \begin{tikzpicture}[vertex/.style={circle, thin, fill=blue!30, draw, inner sep=0, minimum size=5pt}]
        \node (00) [vertex, label=above left:{\(v_{0}\)}] at (-1.25, 0) {};
        \node (01) [vertex, label=above left:{\(v_{1}\)}] at (0,  0) {};
        \node (02) [vertex, label=above left:{\(v_{2}\)}] at (1.25,  0) {};

        \node (10) [vertex, label=above left:{\(v_{3}\)}] at (-1.25, -1.25) {};
        \node (11) [vertex, label=above left:{\(v_{4}\)}] at (0,  -1.25) {};
        \node (12) [vertex, label=above left:{\(v_{5}\)}] at (1.25,  -1.25) {};

@@

    A classical realization \(c\) for this arrangement is
  \begin{align*}
    c(v_{0}) = c(v_{1}) = c(v_{3}) = c(v_{4}) &= +1 \\
    c(v_{2}) = c(v_{5}) &= -1.
  \end{align*}
\end{tcbexample}

@@

  By repeatedly contracting edges, eventually we reach a graph with a
  single vertex.  Due to how we defined the label of each new vertex
  (i.e.\ as the product of the labels of the vertices being merged),
  it follows that the label of the last vertex equals the product of
  the labels of all the original vertices, and therefore equals the
  parity of the arrangement. The remaining vertex has multiple
  self-loops, two for each measurement \(M\) ascribed to the original
  edges.  Since \(M^{2} = I\), going counterclockwise we get a
  sequence \(M_{0}^{2} \ldots M_{\vert E'\vert -1}^{2} = I\). But, we
  saw that contraction preserves the third property, which says that
  the observables around a vertex multiply to the identity times the
  vertex label. Thus, the label of the final vertex is \(1\) and,
  consequently, the parity of the arrangement is \(+1\). \qed
\end{proof}

\begin{tcbexample}{Planar Graph \(\implies\) Not Magic}{planar-nmagic}
  Consider the arrangement depicted in Figure~\ref{fig:class-re} and
  assume there exists a quantum realization that assigns the
  observables \(M_{i}\) to each vertex \(v_{i}\). Suppose that we do
  not know that the parity of this arrangement is \(+1\). We are going
  to follow the strategy described above, to show that its parity must
  be \(+1\). Figure~\ref{fig:class-re-inters} shows the intersection graph
  that corresponds to that arrangement.

  \begin{figure}[H]
    \begin{center}
      \begin{tikzpicture}[vertex/.style={circle, thin, fill=blue!30, draw, inner sep=0, minimum size=5pt}]
        \node (line0) [vertex, label=above:{\(e_{0}\)}] at (0, 0) {};
        \node (line1) [vertex, label=below:{\(e_{1}\)}] at (0, -3) {};
        \node (col0) [vertex, label=left:{\(e_{2}\)}] at (-1.5, -1.5) {};
        \node (col1) [vertex, label=right:{\(e_{3}\)}] at (0, -1.5) {};
        \node (col2) [vertex, label=right:{\(e_{4}\)}] at (1.5, -1.5) {};
        \path[]
        (line0) edge node[sloped, above] {\(M_{0}\)} (col0)
        (line0) edge node[midway, fill=gray!15] {\(M_{1}\)} (col1)
        (line0) edge node[sloped, above] {\(M_{2}\)} (col2)
        (line1) edge node[sloped, below] {\(M_{3}\)} (col0)
        (line1) edge node[midway, fill=gray!15] {\(M_{4}\)} (col1)
        (line1) edge node[sloped, below] {\(M_{5}\)} (col2)
        ;
      \end{tikzpicture}
    \end{center}
    \caption{Intersection graph corresponding to the arrangement seen
      in Figure~\ref{fig:class-re}.\label{fig:class-re-inters}}
  \end{figure}

  Let us contract the edge labelled by \(M_{0}\) (recall that it
  corresponds to the vertex \(v_{0}\) in the arrangement). The
  resulting graph is depicted in Figure~\ref{fig:contract0}.  Notice
  that \(M_{1}M_{2}M_{0} = \ell(e_{0}) I\) and
  \(M_{0}M_{3} = \ell(e_{2}) I\). After contraction, we have
  a new vertex \(e'_{0}\), with \(\ell(e'_{0}) = \ell(e_{0})\ell(e_{2})\)
  and \(M_{1}M_{2}M_{0}^{2}M_{3} = \ell(e_{0})\ell(e_{2}) I = \ell(e'_{0}) I\).

    \begin{figure}[H]
    \begin{center}
      \begin{tikzpicture}[
        vertex/.style={circle, thin, fill=blue!30, draw, inner sep=0, minimum size=5pt},
        every loop/.style={}
        ]
        \node (line1) [vertex, label=below:{\(e_{1}\)}] at (0, -3) {};
        \node (e'0) [vertex, label=left:{\(e'_{0}\)}] at (-1.5, -1.5) {};
        \node (col1) [vertex, label=right:{\(e_{3}\)}] at (0, -1.5) {};
        \node (col2) [vertex, label=right:{\(e_{4}\)}] at (1.5, -1.5) {};
        \path[]
        (e'0) edge[loop above, in=120, out=60, min distance=5mm] node {\(M_{0}^{2}\)} (e'0)
        (e'0) edge node[midway, fill=gray!15] {\(M_{1}\)} (col1)
        (e'0) edge[bend left=45] node[sloped, above] {\(M_{2}\)} (col2)
        (line1) edge node[sloped, below] {\(M_{3}\)} (e'0)
        (line1) edge node[midway, fill=gray!15] {\(M_{4}\)} (col1)
        (line1) edge node[sloped, below] {\(M_{5}\)} (col2)
        ;
      \end{tikzpicture}
    \end{center}
    \caption{Graph that results from the contraction of the edge from
      \(e_{0}\) to \(e_{2}\) in Figure~\ref{fig:class-re-inters}.\label{fig:contract0}}
  \end{figure}

  Let us now contract the edge labelled by \(M_{5}\). The resulting graph
  is shown in Figure~\ref{fig:contract5}.

  \begin{figure}[H]
    \begin{center}
      \begin{tikzpicture}[
        vertex/.style={circle, thin, fill=blue!30, draw, inner sep=0, minimum size=5pt},
        every loop/.style={}
        ]
        \node (e'0) [vertex, label=left:{\(e'_{0}\)}] at (-1.5, -1.5) {};
        \node (col1) [vertex, label=below:{\(e_{3}\)}] at (0, -1.5) {};
        \node (e'1) [vertex, label=right:{\(e'_{1}\)}] at (1.5, -1.5) {};
        \path[]
        (e'0) edge[loop above, in=120, out=60, min distance=5mm] node {\(M_{0}^{2}\)} (e'0)
        (e'0) edge node[midway, fill=gray!15] {\(M_{1}\)} (col1)
        (e'0) edge[bend left=45] node[sloped, above] {\(M_{2}\)} (col2)
        (e'0) edge[bend right=45] node[sloped, below] {\(M_{3}\)} (e'1)
        (e'1) edge node[midway, fill=gray!15] {\(M_{4}\)} (col1)
        (e'1) edge[loop above, in=120, out=60, min distance=5mm] node {\(M_{5}^{2}\)} (e'1)
        ;
      \end{tikzpicture}
    \end{center}
    \caption{Graph that results from the contraction of the edge from
      \(e_{1}\) to \(e_{4}\) in Figure~\ref{fig:contract0}.\label{fig:contract5}}
  \end{figure}

  If we keep contracting edges, eventually we reach a graph with a
  single node, whose observables going counterclockwise are
  \(\prod_{i = 0}^{5} M_{i}^{2} = I\). And, because contracting edges
  preserve the cyclic product property, the product of the observables is the
  identity times the label of the last vertex. Hence, the parity of
  this arrangement is \(+1\) (as we already knew).
\end{tcbexample}

\printbibliography{}
\end{document}
